A novel distributed nature-inspired algorithm for solving optimization problems\thanks{Supported by TECNOLÓGICO NACIONAL DE MÉXICO (TecNM).}

### 2021-09-28 paper.tex
Draft Update
Supported Organization Update
Work in progress
Working on the experiment results.

\begin{document}
%
\title{A novel distributed nature-inspired algorithm for solving optimization problems\thanks{Supported by organization x.}}
% (Biological) Life-Cycle Algorithm: A novel distributed nature-inspired algorithm for solving optimization problems
%
\titlerunning{Life-Cycle Algorithm}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{J. C. Felix-Saul \inst{1} \and
Mario García Valdez\inst{1} \and
Juan J. Merelo Guervós\inst{2}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Department of Graduate Studies, Tijuana Institute of Technology, Tijuana, Mexico \and
Department of Computer Architecture and Technology, Universidad de Granada, Granada, Spain}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in 150--250 words.

Several bio-inspired algorithms use population evolution as analogies of
nature. In this paper, we present an algorithm inspired by the biological
life-cycle of animal species, which consists of several stages: birth, growth,
reproduction, and death. As in nature, we intend to execute all these stages in
parallel and asynchronously on a population that evolves constantly. From the
ground up, we designed the algorithm as a cloud-native solution using the cloud
available resources to divide the processing workload, among several computers
or running the algorithm as a cloud service. The algorithm works concurrently
and asynchronously on a constantly evolving population, using different
computers (or containers) independently, eliminating waiting times between
processes. This algorithm seeks to imitate the natural life cycle, where new
individuals are born at any moment and mature over time, where they age and
suffer mutations throughout their lives. In reproduction, couples match by
mutual attraction, where they may have offspring. Death can happen to everyone:
from a newborn to an aged adult, where the individual's fitness will impact
their longevity. As a proof-of-concept, we implemented the algorithm with
Docker containers by solving the OneMax problem comparing it with a basic
(sequential) GA algorithm, where it showed favorable and promising results.

@@

Bio-inspired algorithms have been very successful when used to solve complex
optimization problems, but as their complexity increases so does the computing
power required. One strategy to address this processing need is to use
distributed computing, or the resources available in the cloud to help find the
solution. This strategy gives us elasticity to increase (or reduce) the
computing power to achieve a balance according to the nature of the problem.

@@

Similar to nature, we identified what most species have in common in the life
cycle. Our algorithm works on a constantly evolving population, experiencing
the stages that all living beings undergo: being born, growing up, reproducing,
and dying, where each of these stages will be processes (of our algorithm) that
randomly affect the population individuals, emulating the organic way.

@@

From the ground up, we designed the algorithm as a cloud-native solution using
the cloud available resources to divide the processing workload, among several
computers or running the algorithm as a cloud service. [The algorithm works
concurrently and asynchronously on a constantly evolving population, using
different computers (or containers) independently, eliminating waiting times
between processes.]

In this paper, we present an algorithm inspired by the biological life-cycle of
animal species, which consists of several stages: birth, growth, reproduction,
and death. As in nature, we intend to execute all these stages in parallel and
asynchronously on a population that evolves constantly. We show the LifeCycle 
Algorithm Proposal Model in Fig.~\ref{fig2}.

@@

\begin{figure}
    \includegraphics[width=\textwidth]{img/fig2_proposal.pdf}
    \caption{LifeCycle Algorithm Proposal Model.} \label{fig2}
    \end{figure}

@@

The attraction of a couple will depend on fitness: the better individual's
fitness, the more attractive it will be, making it easier to find mating
matches. Not all couples will be compatible, so reproduction will not always be
possible, but the problem arises: how to quantify the attraction between two
individuals?. This process will be taking random pairs of individuals to
evaluate their attraction as a couple, to try to breed; when the gestation is
successful, a new pair of individuals will be born (as the offspring).

@@

The death stage represents the challenges and adversities that life presents
(to overcome). This process evaluates the individual resistance to survive in
the environment. The better fitness the individual has will increase its
chances of survival. As time progresses, the demands of nature will also
increase, pushing for only the best individuals to survive.

@@

As a proof-of-concept, we implemented the algorithm with Docker containers by
solving the OneMax problem to compare it with a traditional (sequential) GA
algorithm using the DEAP library. The OneMax problem uses a genetic algorithm
to evolve a population of randomly generated individuals with zeros and ones,
and it stops until a solution of only ones is found.

In this experiment, we needed to match or balance the experimentation
parameters according to the specifications of the algorithm used by the DEAP
library. This is to be able to verify if our algorithm would also converge on
the solutions in a similar number of evaluations and execution time.

\subsection{Experimental Setup}
    Table~\ref{tab1} shows the configuration for the OneMax DEAP version, while Table~\ref{tab2} shows the LifeCycle Algorithm setup.

    \begin{table}[]
        \centering
        \caption{OneMax DEAP configuration.}\label{tab1}
        \begin{tabular}{|l|l|}
        \hline
        \multicolumn{2}{|l|}{OneMax   (DEAP)} \\ \hline
        Population & 60 \\ \hline
        Max   Generation & 20 \\ \hline
        Stagnation & Off \\ \hline
        Chromosome   Length & 20 \\ \hline
        Target   Fitness & 20 \\ \hline
        Crossover   Rate & 100 \\ \hline
        Mutation   Rate & 7 \\ \hline
        Tourney   Reprod. Rate & 50 \\ \hline
        Tourney   Sample & 3 \\ \hline
        \end{tabular}
        \end{table}

    \begin{table}[]
        \centering        
        \caption{LifeCycle Algorithm configuration.}\label{tab2}
        \begin{tabular}{|l|l|}
        \hline
        \multicolumn{2}{|l|}{LifeCycle   Algorithm} \\ \hline
        Population & 60 \\ \hline
        Max   Generation & 20 \\ \hline
        Stagnation & Off \\ \hline
        Chromosome   Length & 20 \\ \hline
        Target   Fitness & 20 \\ \hline
        Crossover   Rate & 100 \\ \hline
        Mutation   Rate & 7 \\ \hline
        Tourney   Reprod. Rate & 50 \\ \hline
        Tourney   Sample & 4 \\ \hline
        Max   Age & 80 \\ \hline
        Open   Reprod. Rate & 5 \\ \hline
        Closed   Reprod. Rate & 45 \\ \hline
        Fertility   Rate & 100 \\ \hline
        Base   Approval & 80 \\ \hline
        Goal   Approval & 100 \\ \hline
        Max   Wait Time & 2 \\ \hline
        \end{tabular}
        \end{table}

\subsection{Results}
    Table~\ref{tab3} shows the OneMax DEAP experiment results, while Table~\ref{tab4} shows the LifeCycle Algorithm experiment results.

>>>>

\begin{document}
%
\title{A novel distributed nature-inspired algorithm for solving optimization problems\thanks{Supported by TECNOLÓGICO NACIONAL DE MÉXICO (TecNM).}}
% (Biological) Life-Cycle Algorithm: A novel distributed nature-inspired algorithm for solving optimization problems
%
\titlerunning{Life-Cycle Algorithm}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{J. C. Felix-Saul \inst{1} \and
Mario García Valdez\inst{1} \and
Juan J. Merelo Guervós\inst{2}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Department of Graduate Studies, Tijuana Institute of Technology, Tijuana, Mexico \and
Department of Computer Architecture and Technology, Universidad de Granada, Granada, Spain}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in 150--250 words.

Several bio-inspired algorithms use population evolution as analogies of
na-ture. In this paper, we present an algorithm inspired by the biological
Life-Cycle of animal species, which consists of several stages: birth, growth,
re-production, and death. As in nature, we intend to execute all these stages
in parallel and asynchronously on a population that evolves constantly. From
the ground up, we designed the algorithm as a cloud-native solution using the
cloud available resources to divide the processing workload, among several
computers or running the algorithm as a cloud service. The algorithm works
concurrently and asynchronously on a constantly evolving population, using
different computers (or containers) independently, eliminating waiting times
between processes. This algorithm seeks to imitate the natural life cycle,
where new individuals are born at any moment and mature over time, where they
age and suffer mutations throughout their lives. In reproduction, couples match
by mutual attraction, where they may have offspring. Death can happen to
everyone: from a newborn to an aged adult, where the individual's fitness will
impact their longevity. As a proof-of-concept, we implemented the algo-rithm
with Docker containers by solving the OneMax problem comparing it with a basic
(sequential) GA algorithm, where it showed favorable and prom-ising results.

@@

Bio-inspired algorithms have been very successful when used to solve complex
optimization problems, but as their complexity increases so does the computing
power required. One strategy to address this processing need is to use
distributed computing, or the resources available in the cloud to help find the
solution. This strategy gives us elasticity to adjust (increase or reduce) the
computing power to achieve a balance according to the nature of the problem.

@@

Like nature, we identified what most species have in common in the life cycle.
Our algorithm works on a constantly evolving population, experiencing the
stages that all living beings undergo: being born, growing up, reproducing, and
dying, where each of these stages will be processes (of our algorithm) that
randomly affect the population individuals, emulating the organic way.

@@

In this paper, we present an algorithm inspired by the biological Life-Cycle of
animal species, which consists of several stages: birth, growth, reproduction,
and death. As in nature, we intend to execute all these stages in parallel and
asynchronously on a population that evolves constantly.

From the ground up, we designed the algorithm as a cloud-native solution using
the cloud available resources to divide the processing workload, among several
computers or running the algorithm as a cloud service. The algorithm works
concurrently and asynchronously on a constantly evolving population, using
different computers (or containers) independently, theoretically eliminating
waiting times between processes.

@@

\begin{figure}
    \includegraphics[width=\textwidth]{img/fig2_proposal.pdf}
    \caption{Algorithm model inspired by the animal species' biological Life-Cycle.} \label{fig2}
    \end{figure}

@@

The attraction of a couple will depend on fitness: the better individual's
fitness, the more attractive it will be, making it easier to find mating
matches. This process will be taking random pairs of individuals to evaluate
their attraction as a couple, to try to breed; when the gestation is
successful, a new pair of individuals will be born (as the offspring). Not all
couples will be compatible, so reproduction will not always be possible, but
the problem arises: how to quantify the attraction between two individuals?

We could have used several strategies to evaluate this attraction. Considering
this algorithm takes a similar focus as the study of bacteria growth in
microbiology, where we can observe and analyze the evolution of a population
over time. As in nature' most species, the bigger a specimen is, the more
attractive it is to its mating candidates. To be consistent with both ideas in
our algorithm, we use the equation of Newton's Law of Universal Gravitation (1)
to calculate the attraction between two individuals, where we visualize the
individual fitness as its mass when using the equation. In previous work,
Newton's Law of Universal Gravitation has shown success in helping to solve
optimization problems.

\begin{figure}
    \includegraphics[width=\textwidth]{img/fig3_newtonslaw.pdf}
    \caption{Newton's Law of Universal Gravitation.} \label{fig3}
    \end{figure}

Where:
•	F is the gravitational force acting between two objects
•	m1 and m2 are the masses of the objects
•	r is the distance between the centers of their masses
•	G is the gravitational constant: 6.67430(15)×10−11 m3⋅kg−1⋅s−2

@@

The death stage represents the challenges and adversities that life presents to
overcome. This process evaluates the individual resistance to survive in the
environment. The better fitness the individual has will increase its chances of
survival. As time progresses, the demands of nature will also increase, pushing
for only the best individuals to survive.

\subsection{Life-Cycle algorithm flowchart}

The figure 4 describes the general flowchart for the Life-Cycle algorithm.

\begin{figure}
    \includegraphics[width=\textwidth]{img/fig4_flowchart.pdf}
    \caption{Life-Cycle algorithm general flowchart.} \label{fig4}
    \end{figure}

@@

As a proof-of-concept, we implemented the algorithm with Docker containers by
solving the OneMax problem to compare it with a traditional GA algorithm using
the DEAP (Distributed Evolutionary Algorithms in Python) library. The OneMax
problem uses a genetic algorithm to evolve a population of randomly generated
individuals with zeros and ones, and it stops until a solution of only ones is
found.

\subsection{Experimental Setup}

\begin{figure}
    \includegraphics[width=\textwidth]{img/fig5_processes_containers .pdf}
    \caption{Life-Cycle algorithm processes in containers comparison.} \label{fig5}
    \end{figure}

One strategic difference in our algorithm implementation is the reproduction
process, which is flexible to work in parallel with multiple couple selection
methods, for example, tournament selection, random selection (closed
reproduction), and couple match by creating a new mating individual (open
reproduction). We use many combinations of the containerized processes, up to
the minimum number required to obtain good results. Figure 4 shows a comparison
of only two alternatives for the implementation of the algorithm. Our
experimentation started with a configuration of ten processes and ended in
five, shown on the left and right sides of figure 4, respectively.

\subsection{Experiment Configuration}

In this experiment, we needed to match or balance the experimentation
parameters according to the specifications of the algorithm used by the DEAP
library. This is to be able to verify if our algorithm would also converge on
the solutions in a similar number of evaluations and execution time.
Table~\ref{tab1} shows the OneMax initial configuration for DEAP and Life-Cycle
experiment.

\begin{table}[]
    \centering        
    \caption{OneMax initial configuration for DEAP and Life-Cycle experiment.}\label{tab1}
    \begin{tabular}{lll}
    \textbf{Configuration} & \textbf{DEAP} & \textbf{Life-Cycle} \\
    Population & 60 & 60 \\
    Max Generation & 20 & 20 \\
    Stagnation & Off & 10 \\
    Chromosome Length & 20 & 20 \\
    Target Fitness & 20 & 20 \\
    Crossover Rate & 100 & 100 \\
    Mutation Rate & 7 & 7 \\
    Tournament Rate & 100 & 50 \\
    Tournament Sample & 3 & 4 \\
    Open Reproduction Rate & NA & 5 \\
    Closed Reproduction Rate & NA & 45 \\
    Max Age & NA & 80 \\
    Base Approval & NA & 80 \\
    Goal Approval & NA & 100
    \end{tabular}
    \end{table}

\subsection{Experiment Results}

For each experiment, we ran 60 independent executions per algorithm and
recorded the following results: Last Generation, Total Evaluations, Time
(seconds), Evaluations/second.

•	Last Gen.: The generation that found the solution.
•	Eval.: The total number of evaluations the algorithm executed. 
•	Time (sec): Total elapsed or wall clock time, in seconds. 
•	Eval/sec: The calculated rate of evaluations per second.

Experiment 1: The goal of our first OneMax experiment was to make sure our
Life-Cycle algorithm worked as expected, converging on the solution. For this
experiment, we intentionally used delays (on the tenth of a second magnitude)
to follow the Life-Cycle algorithm behavior on the console output. For this
experiment, we compared the DEAP algorithm versus the Life-Cycle using ten
processes (on containers) where we show the summarized results in Table 2, and
Figure 5 shows their Box and Whisker chart.

Experiment 2: The goal of our second experiment was to confirm the Life-Cycle
algorithm continued working as expected, converging on the solution. For this
experiment, we reduced the time used on delays (now on the thousands of a
second magnitude) to follow the Life-Cycle algorithm behavior. For this
experiment, we only used tournament selection for the reproduction, on the
first configuration running the Life-Cycle on ten processes, versus the second
configuration where we reduced the processes to the minimum basic 5. We show
the summarized results in Table 3, and Figure 6 shows their Box and Whisker
chart.

Experiment 3: The goal of our third experiment was to follow and study the
behavior of the Life-Cycle algorithm that continued converging on the solution.
For this experiment, we remained using minimal delays (thousands of a second
magnitude). For this experiment, we only used tournament selection for the
reproduction, on the first configuration running the Life-Cycle on six
processes, two of whom was the Death stage, versus the second configuration
where we reduced the processes to the minimum (five) but increasing the (Death)
goal approval range to 115. We show the summarized results in Table 4, and
Figure 7 shows their Box and Whisker chart.

Experiment 4: The goal of our fourth and last experiment was to follow and
study the behavior of the Life-Cycle algorithm that continued converging on the
solution. For this experiment, we remained using minimal delays (thousands of a
second magnitude). For this experiment, we compared the OneMax DEAP
implementation versus the Life-Cycle algorithm, only using tournament selection
for the reproduction, with the Life-Cycle configuration running on the minimum
(five) processes but increasing the (Death) goal approval range to 125. We show
the summarized results in Table 5, and Figure 8 shows their Box and Whisker
chart.

    Table~\ref{tab3} shows the OneMax DEAP experiment results, while Table~\ref{tab4} shows the LifeCycle Algorithm experiment results.
